### 2013-07-27 doc/taxize_withcode_plos.tex
incorporated changes from karthik and naim, fixes #170, should be close to submit

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

@@

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}
Evolution by natural selection has led to a hierarchical relationship among all living organisms.  Thus, species are categorized using a taxonomic hierarchy, starting with the binomial species name (e.g, \emph{Homo sapiens}), moving up to genus (\emph{Homo}), then family (\emph{Hominidae}), and on up to Domain (\emph{Eukarya}). Biologists, whether studying organisms at the cell, organismal, or community level, can put their study taxa into taxonomic context, allowing them to know close and distant relatives, find relevant literature, and more.

The use of taxonomic names is, unfortunately, not straightforward. Taxonomic names often vary due to name changes at the generic or specific levels, lumping or splitting lower taxa (genera, species) among higher taxa (families), and name spelling changes. For example, a study found that a compilation of 308,000 plant observations from 51 digitized herbarium records had 22,100 unique taxon names, of which only ~13,000 were accepted names \cite{weiser2007,boyle2013}. In addition, there is no one authoritative taxonomic names source for all taxa - although, there are taxon specific sources that are used by many scientists. Different sources (e.g., uBio, Tropicos, ITIS) may use different accepted names for the same taxon. For example, while the Integrated Taxonomic Information Service (ITIS) has \emph{Helianthus x glaucus} as an accepted name, The Plant List \cite{theplantlist} has that name as unresolved. But \emph{Helianthus glaucus} is an accepted name in The Plant List, while ITIS does not list this name.

@@

Taxonomic data is getting easier to obtain through the web (e.g., \cite{eol}). However, there are a number of good reasons to obtain taxonomic information programatically rather than through a web interface. First, if you have more than a few names to lookup on a website, it can take quite a long time to enter each name, get data, and repeat for each species. Programatically getting taxonomic names solves the problem by looping over a list of names. In addition, doing taxonomic searching, etc. becomes reproducible. With increasing reports of irreproducibility in science \cite{stodden2010,zimmer2012}, it is extremely important to make science workflows repeatable. Science workflows can now easily incorporate text, code, and images in a single executable document \cite{yihui2013}.

The R language is a widely used language by biologists, and now has over 4,000 packages on the Comprehensive R Archive Network (CRAN) and more than 2,500 packages on other repositories to extend R. R is great for manipulating, visualizing and fitting statistical models to data. \cite{gentleman_bioconductor:_2004} gives a detailed discussion of advantages of R in computational biology. Getting data from the web will be increasingly common as more and more data gets moved to the cloud. Therefore there is a need to get data from the web directly into R. Increasingly, data is available from the web via application programming interfaces (API). These allow computers to talk to one another using code that is not human readable, but is machine readable. Web APIs often define a number of methods that allow users to search for a species name, or retrieve the synonyms for a species name, for example. A further advantage of APIs is that they are language agnostic, meaning that data can be consumed in almost any computing context, allowing users to interact with the web API without having to know the details of the code. Moreover data can be accessed from every computer, whereas for example an Excel file can only be opened in a few programs.

The goal of taxize, an R package in development, is to make many use cases having to do with retrieving and resolving taxonomic names easy and reproducible. In taxize, we have written a suite of R functions that interact with many taxonomic data sources via their web APIs (Table~\ref{tab:a}). The interface to each function is usually a simple list of species names, just as a user would do with a web API. Therefore, we hope moving from a web to R interface for taxonomic names will be relatively seamless (if one is already nominally familiar with R).

Here, we justify the need for taxize, discuss our data sources, and run through a suite of use cases to demonstrate the variety of ways that users can interact with taxize.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{install.packages}(\hlstr{"taxize"})
\hlkwd{library}(taxize)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
temp <- \hlkwd{gnr_resolve}(names = \hlkwd{c}(\hlstr{"Helianthos annus"}, \hlstr{"Homo saapiens"}), returndf = TRUE)
temp[, -\hlkwd{c}(1, 4)]
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
mynames <- \hlkwd{c}(\hlstr{"Helianthus annuus"}, \hlstr{"Pinus contort"}, \hlstr{"Poa anua"}, \hlstr{"Abis magnifica"}, 
    \hlstr{"Rosa california"}, \hlstr{"Festuca arundinace"}, \hlstr{"Sorbus occidentalos"}, \hlstr{"Madia sateva"})
\hlkwd{tnrs}(query = mynames)[, -\hlkwd{c}(5:7)]
\end{alltt}
\begin{verbatim}
        submittedName        acceptedName    sourceId score
7   Helianthus annuus   Helianthus annuus iPlant_TNRS  1.00
4       Pinus contort      Pinus contorta iPlant_TNRS  0.98
5            Poa anua           Poa annua iPlant_TNRS  0.96
3      Abis magnifica     Abies magnifica iPlant_TNRS  0.96
8     Rosa california    Rosa californica iPlant_TNRS  0.99
2  Festuca arundinace Festuca arundinacea iPlant_TNRS  0.99
1 Sorbus occidentalos Sorbus occidentalis iPlant_TNRS  0.99
6        Madia sateva        Madia sativa iPlant_TNRS  0.97
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
specieslist <- \hlkwd{c}(\hlstr{"Abies procera"}, \hlstr{"Pinus contorta"})
\hlkwd{classification}(specieslist, db = \hlstr{"itis"})
\end{alltt}
\begin{verbatim}
$`Abies procera`
        rankName       taxonName    tsn
1        Kingdom         Plantae 202422
2     Subkingdom  Viridaeplantae 846492
3   Infrakingdom    Streptophyta 846494
4       Division    Tracheophyta 846496
5    Subdivision Spermatophytina 846504
6  Infradivision    Gymnospermae 846506
7          Class       Pinopsida 500009
8          Order         Pinales 500028
9         Family        Pinaceae  18030
10         Genus           Abies  18031
11       Species   Abies procera 181835

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{tax_name}(query = \hlstr{"Helianthus annuus"}, get = \hlstr{"family"}, db = \hlstr{"itis"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\begin{alltt}
\hlkwd{tax_name}(query = \hlstr{"Helianthus annuus"}, get = \hlstr{"family"}, db = \hlstr{"ncbi"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{get_tsn}(searchterm = \hlstr{"Heliastes"}, searchtype = \hlstr{"sciname"})
\end{alltt}
\begin{verbatim}
           combinedname    tsn
1     Heliastes bicolor 615238
2   Heliastes chrysurus 615250
3     Heliastes cinctus 615573
4  Heliastes dimidiatus 615257
5  Heliastes hypsilepis 615273
6 Heliastes immaculatus 615639
7 Heliastes opercularis 615300
8      Heliastes ovalis 615301
 1 
NA 
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

In another example, you can pass in a long list of taxonomic names:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
splist <- \hlkwd{c}(\hlstr{"annona cherimola"}, \hlstr{"annona muricata"}, \hlstr{"quercus robur"}, \hlstr{"shorea robusta"}, 
    \hlstr{"pandanus patina"}, \hlstr{"oryza sativa"}, \hlstr{"durio zibethinus"})
\hlkwd{get_tsn}(searchterm = splist, searchtype = \hlstr{"sciname"})
\end{alltt}
\begin{verbatim}
[1] "506198" "18098"  "19405"  "506787" "507376" "41976"  "506099"
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{get_uid}(sciname = \hlkwd{c}(\hlstr{"Chironomus riparius"}, \hlstr{"aaa vva"}))
\end{alltt}
\begin{verbatim}
[1] "315576" NA      
attr(,"class")
[1] "uid"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
taxa <- \hlkwd{c}(\hlstr{"Poa annua"}, \hlstr{"Abies procera"}, \hlstr{"Helianthus annuus"})
tree <- \hlkwd{phylomatic_tree}(taxa = taxa)
tree$tip.label <- \hlkwd{capwords}(tree$tip.label)
\hlkwd{plot}(tree, cex = 1)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{col_downstream}(name = \hlstr{"Apis"}, downto = \hlstr{"Species"})[[1]]
\end{alltt}
\begin{verbatim}
  childtaxa_id     childtaxa_name childtaxa_rank
1      6971712 Apis andreniformis        Species
2      6971713        Apis cerana        Species
3      6971714       Apis dorsata        Species
4      6971715        Apis florea        Species
5      6971716 Apis koschevnikovi        Species
6      6845885     Apis mellifera        Species
7      6971717   Apis nigrocincta        Species
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
ia <- \hlkwd{iucn_summary}(\hlkwd{c}(\hlstr{"Panthera uncia"}, \hlstr{"Lynx lynx"}))
\hlkwd{iucn_status}(ia)
\end{alltt}
\begin{verbatim}
Panthera uncia      Lynx lynx 
          "EN"           "LC" 
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out <- \hlkwd{get_genes_avail}(taxon_name = \hlstr{"Umbra limi"}, seqrange = \hlstr{"1:2000"}, getrelated = FALSE)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out[\hlkwd{grep}(\hlstr{"RAG1"}, out$genesavail, ignore.case = TRUE), ]
\end{alltt}
\begin{verbatim}
        spused length
413 Umbra limi    732
427 Umbra limi    959
434 Umbra limi   1631
                                                                            genesavail
413 isolate UlimA recombinase activating protein 1 (rag1) gene, exon 3 and partial cds
427           recombination-activating protein 1 (RAG1) gene, intron 2 and partial cds
434                        recombination-activating protein 1 (RAG1) gene, partial cds
    access_num       ids
413   JX190826 394772608
427   AY459526  45479841
434   AY380548  38858304
\end{verbatim}
\end{kframe}
\end{knitrout}

It turns out that there are XX different unique records found. However, this doesn't mean that there are XX different genes found as the API does not provide metadata to classify genes. However, at the end of the example, we showed that you can use regular expressions (e.g., via \emph{grep}) to search for the gene of interest.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
A <- \hlstr{"gammarus roeseli"}

B1 <- \hlstr{"gammarus roeseli"}
B2 <- \hlstr{"gammarus"}
B3 <- \hlstr{"gammaridae"}

A_clas <- \hlkwd{classification}(A, db = \hlstr{'ncbi'})
B1_clas <- \hlkwd{classification}(B1, db = \hlstr{'ncbi'})
B2_clas <- \hlkwd{classification}(B2, db = \hlstr{'ncbi'})
B3_clas <- \hlkwd{classification}(B3, db = \hlstr{'ncbi'})

B1[\hlkwd{match}(A, B1)]
\end{alltt}
\begin{verbatim}
[1] "gammarus roeseli"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlkwd{tolower}(A_clas[[1]]$ScientificName) %in% B2]
\end{alltt}
\begin{verbatim}
[1] "genus"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlkwd{tolower}(A_clas[[1]]$ScientificName) %in% B3]
\end{alltt}
\begin{verbatim}
[1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{data}(dune, package = \hlstr{'vegan'})
df <- dune[ , 1:5]
\hlkwd{colnames}(df) <- \hlkwd{c}(\hlstr{"Bellis perennis"}, \hlstr{"Empetrum nigrum"}, \hlstr{"Juncus bufonius"}, 
                  \hlstr{"Juncus articulatus"}, \hlstr{"xxx"})
\hlkwd{head}(df)
\end{alltt}
\begin{verbatim}
   Bellis perennis Empetrum nigrum Juncus bufonius Juncus articulatus xxx
2                3               0               0                  0   0
13               0               0               3                  0   0
4                2               0               0                  0   0
16               0               0               0                  3   0
6                0               0               0                  0   0
1                0               0               0                  0   0
\end{verbatim}
\begin{alltt}
agg <- \hlkwd{tax_agg}(df, rank = \hlstr{'family'}, db = \hlstr{'ncbi'})
agg
\end{alltt}
\begin{verbatim}

>>>>

\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
\newcommand{\hlslot}[1]{\textit{#1}}%
\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%

@@

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
% \section*{Author Summary}

\section*{Introduction}
Evolution by natural selection has led to a hierarchical relationship among all living organisms.  Thus, species are categorized using a taxonomic hierarchy, starting with the binomial species name (e.g, \emph{Homo sapiens}), moving up to genus (\emph{Homo}), then family (\emph{Hominidae}), and on up to Domain (\emph{Eukarya}). Biologists, whether studying organisms at the cell, organismal, or community level, can put their study objects into taxonomic context, allowing them to know close and distant relatives, find relevant literature, and more.

The use of taxonomic names is, unfortunately, not straightforward. Taxonomic names often vary due to name revisions at the generic or specific levels, lumping or splitting lower taxa (genera, species) among higher taxa (families), and name spelling changes. For example, a study found that a compilation of 308,000 plant observations from 51 digitized herbarium records had 22,100 unique taxon names, of which only ~13,000 were accepted names \cite{weiser2007,boyle2013}. In addition, there is no one authoritative taxonomic names source for all taxa - although, there are taxon specific sources that are used by many scientists. Different sources (e.g., uBio, Tropicos, ITIS) may use different accepted names for the same taxon. For example, while the Integrated Taxonomic Information Service (ITIS) has \emph{Helianthus x glaucus} as an accepted name, The Plant List \cite{theplantlist} has that name as unresolved. But \emph{Helianthus glaucus} is an accepted name in The Plant List, while ITIS does not list this name.

@@

Taxonomic data is getting easier to obtain through the web (e.g., \cite{eol}). However, there are a number of good reasons to obtain taxonomic information programatically rather than through a web interface. First, if you have more than a few names to lookup on a website, it can take quite a long time to enter each name, get data, and repeat for each species. Programatically getting taxonomic names solves the problem by looping over a list of names. In addition, doing taxonomic searching, etc. becomes reproducible. With increasing reports of irreproducibility in science \cite{stodden2010,zimmer2012}, it is extremely important to make science workflows repeatable. Science workflows can now easily incorporate text, code, and images in a single executable document \cite{yihui2013}. Reproducible documents should become mainstream in biology to avoid mistakes, and make collaboration easier.

The R language is a widely used language by biologists, and now has over 5,000 packages on the Comprehensive R Archive Network (CRAN) to extend R. R is great for manipulating, visualizing and fitting statistical models to data. \cite{gentleman_bioconductor:_2004} gives a detailed discussion of advantages of R in computational biology. Getting data from the web will be increasingly common as more and more data gets moved to the cloud. Therefore there is a need to get data from the web directly into R. Increasingly, data is available from the web via application programming interfaces (API). These allow computers to talk to one another using code that is not human readable, but is machine readable. Web APIs often define a number of methods that allow users to search for a species name, or retrieve the synonyms for a species name, for example. A further advantage of APIs is that they are language agnostic, meaning that data can be consumed in almost any computing context, allowing users to interact with the web API without having to know the details of the code. Moreover data can be accessed from every computer, whereas for example an Excel file can only be opened in a few programs.

The goal of taxize, an R package in development, is to make many use cases having to do with retrieving and resolving taxonomic names easy and reproducible. In taxize, we have written a suite of R functions that interact with many taxonomic data sources via their web APIs (Table~\ref{tab:a}). The interface to each function is usually a simple list of species names, just as a user would do when interacting with a website. Therefore, we hope moving from a web to R interface for taxonomic names will be relatively seamless (if one is already nominally familiar with R).

Here, we justify the need for programmatic taxonomic resolution tools like taxize, discuss our data sources, and run through a suite of use cases to demonstrate the variety of ways that users can use taxize.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{install.packages}(\hlstring{"taxize"})
\hlfunctioncall{library}(taxize)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
temp <- \hlfunctioncall{gnr_resolve}(names = \hlfunctioncall{c}(\hlstring{"Helianthos annus"}, \hlstring{"Homo saapiens"}))
temp[, -\hlfunctioncall{c}(1, 4)]
\end{alltt}
\begin{verbatim}
                 matched_name       data_source_title
1        Helianthus annuus L.       Catalogue of Life
2            Helianthus annus GBIF Taxonomic Backbone
3            Helianthus annus                     EOL
4         Helianthus annus L.                     EOL
5            Helianthus annus           uBio NameBank
6 Homo sapiens Linnaeus, 1758       Catalogue of Life
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
mynames <- \hlfunctioncall{c}(\hlstring{"Helianthus annuus"}, \hlstring{"Pinus contort"}, \hlstring{"Poa anua"}, \hlstring{"Abis magnifica"}, 
    \hlstring{"Rosa california"}, \hlstring{"Festuca arundinace"}, \hlstring{"Sorbus occidentalos"}, \hlstring{"Madia sateva"})
\hlfunctioncall{tnrs}(query = mynames)[, -\hlfunctioncall{c}(5:7)]
\end{alltt}
\begin{verbatim}
         submittedName        acceptedName    sourceId score
9    Helianthus annuus   Helianthus annuus iPlant_TNRS     1
10   Helianthus annuus   Helianthus annuus        NCBI     1
4        Pinus contort      Pinus contorta iPlant_TNRS  0.98
5             Poa anua           Poa annua iPlant_TNRS  0.96
3       Abis magnifica     Abies magnifica iPlant_TNRS  0.96
7      Rosa california    Rosa californica iPlant_TNRS  0.99
8      Rosa california          California        NCBI     1
2   Festuca arundinace Festuca arundinacea iPlant_TNRS  0.99
1  Sorbus occidentalos Sorbus occidentalis iPlant_TNRS  0.99
6         Madia sateva        Madia sativa iPlant_TNRS  0.97
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

Another common use case is when there are many synonyms for a species. In this example, we have three synonyms of the currently accepted name for a species.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
mynames <- \hlfunctioncall{c}(\hlstring{"Helianthus annuus ssp. jaegeri"}, \hlstring{"Helianthus annuus ssp. lenticularis"}, 
    \hlstring{"Helianthus annuus ssp. texanus"}, \hlstring{"Helianthus annuus var. lenticularis"}, 
    \hlstring{"Helianthus annuus var. macrocarpus"}, \hlstring{"Helianthus annuus var. texanus"})
tsn <- \hlfunctioncall{get_tsn}(mynames)
\hlfunctioncall{ldply}(tsn, itis_acceptname)
\end{alltt}
\begin{verbatim}
  submittedTsn      acceptedName acceptedTsn
1       525928 Helianthus annuus       36616
2       525929 Helianthus annuus       36616
3       525930 Helianthus annuus       36616
4       536095 Helianthus annuus       36616
5       536096 Helianthus annuus       36616
6       536097 Helianthus annuus       36616
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
specieslist <- \hlfunctioncall{c}(\hlstring{"Abies procera"}, \hlstring{"Pinus contorta"})
\hlfunctioncall{classification}(specieslist, db = \hlstring{"itis"})
\end{alltt}
\begin{verbatim}
$`Abies procera`
        rankName       taxonName    tsn
1        Kingdom         Plantae 202422
2     Subkingdom  Viridaeplantae 846492
3   Infrakingdom    Streptophyta 846494
4       Division    Tracheophyta 846496
5    Subdivision Spermatophytina 846504
6  Infradivision    Gymnospermae 846506
7          Class       Pinopsida 500009
8          Order         Pinales 500028
9         Family        Pinaceae  18030
10         Genus           Abies  18031
11       Species   Abies procera 181835

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"itis"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\begin{alltt}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"ncbi"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{get_tsn}(searchterm = \hlstring{"Heliastes"}, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}
           combinedname    tsn
1     Heliastes bicolor 615238
2   Heliastes chrysurus 615250
3     Heliastes cinctus 615573
4  Heliastes dimidiatus 615257
5  Heliastes hypsilepis 615273
6 Heliastes immaculatus 615639
7 Heliastes opercularis 615300
8      Heliastes ovalis 615301
 1 
NA 
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

In another example, you can pass in a long character vector of taxonomic names:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
splist <- \hlfunctioncall{c}(\hlstring{"annona cherimola"}, \hlstring{"annona muricata"}, \hlstring{"quercus robur"}, \hlstring{"shorea robusta"}, 
    \hlstring{"pandanus patina"}, \hlstring{"oryza sativa"}, \hlstring{"durio zibethinus"})
\hlfunctioncall{get_tsn}(searchterm = splist, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}
[1] "506198" "18098"  "19405"  "506787" "507376" "41976"  "506099"
attr(,"class")
[1] "tsn"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{get_uid}(sciname = \hlfunctioncall{c}(\hlstring{"Chironomus riparius"}, \hlstring{"aaa vva"}))
\end{alltt}
\begin{verbatim}
[1] "315576" NA      
attr(,"class")
[1] "uid"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
taxa <- \hlfunctioncall{c}(\hlstring{"Poa annua"}, \hlstring{"Abies procera"}, \hlstring{"Helianthus annuus"})
tree <- \hlfunctioncall{phylomatic_tree}(taxa = taxa)
tree$tip.label <- \hlfunctioncall{capwords}(tree$tip.label)
\hlfunctioncall{plot}(tree, cex = 1)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{col_downstream}(name = \hlstring{"Apis"}, downto = \hlstring{"Species"})[[1]]
\end{alltt}
\begin{verbatim}
  childtaxa_id     childtaxa_name childtaxa_rank
1      6971712 Apis andreniformis        Species
2      6971713        Apis cerana        Species
3      6971714       Apis dorsata        Species
4      6971715        Apis florea        Species
5      6971716 Apis koschevnikovi        Species
6      6845885     Apis mellifera        Species
7      6971717   Apis nigrocincta        Species
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
ia <- \hlfunctioncall{iucn_summary}(\hlfunctioncall{c}(\hlstring{"Panthera uncia"}, \hlstring{"Lynx lynx"}))
\hlfunctioncall{iucn_status}(ia)
\end{alltt}
\begin{verbatim}
Panthera uncia      Lynx lynx 
          "EN"           "LC" 
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out <- \hlfunctioncall{get_genes_avail}(taxon_name = \hlstring{"Umbra limi"}, seqrange = \hlstring{"1:2000"}, getrelated = FALSE)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
out[\hlfunctioncall{grep}(\hlstring{"RAG1"}, out$genesavail, ignore.case = TRUE), ]
\end{alltt}
\begin{verbatim}
        spused length
413 Umbra limi    732
427 Umbra limi    959
434 Umbra limi   1631
                                                                            genesavail
413 isolate UlimA recombinase activating protein 1 (rag1) gene, exon 3 and partial cds
427           recombination-activating protein 1 (RAG1) gene, intron 2 and partial cds
434                        recombination-activating protein 1 (RAG1) gene, partial cds
    access_num       ids
413   JX190826 394772608
427   AY459526  45479841
434   AY380548  38858304
\end{verbatim}
\end{kframe}
\end{knitrout}

It turns out that there are 430 different unique records found. However, this doesn't mean that there are 430 different genes found as the API does not provide metadata to classify genes. However, you can use regular expressions (e.g., \emph{grep}) to search for the gene of interest.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
A <- \hlstring{"gammarus roeseli"}

B1 <- \hlstring{"gammarus roeseli"}
B2 <- \hlstring{"gammarus"}
B3 <- \hlstring{"gammaridae"}

A_clas <- \hlfunctioncall{classification}(A, db = \hlstring{'ncbi'})
B1_clas <- \hlfunctioncall{classification}(B1, db = \hlstring{'ncbi'})
B2_clas <- \hlfunctioncall{classification}(B2, db = \hlstring{'ncbi'})
B3_clas <- \hlfunctioncall{classification}(B3, db = \hlstring{'ncbi'})

B1[\hlfunctioncall{match}(A, B1)]
\end{alltt}
\begin{verbatim}
[1] "gammarus roeseli"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlfunctioncall{tolower}(A_clas[[1]]$ScientificName) %in% B2]
\end{alltt}
\begin{verbatim}
[1] "genus"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlfunctioncall{tolower}(A_clas[[1]]$ScientificName) %in% B3]
\end{alltt}
\begin{verbatim}
[1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{data}(dune, package = \hlstring{'vegan'})
df <- dune[ , 1:5]
\hlfunctioncall{colnames}(df) <- \hlfunctioncall{c}(\hlstring{"Bellis perennis"}, \hlstring{"Empetrum nigrum"}, \hlstring{"Juncus bufonius"}, 
                  \hlstring{"Juncus articulatus"}, \hlstring{"xxx"})
\hlfunctioncall{head}(df)
\end{alltt}
\begin{verbatim}
   Bellis perennis Empetrum nigrum Juncus bufonius Juncus articulatus xxx
2                3               0               0                  0   0
13               0               0               3                  0   0
4                2               0               0                  0   0
16               0               0               0                  3   0
6                0               0               0                  0   0
1                0               0               0                  0   0
\end{verbatim}
\begin{alltt}
agg <- \hlfunctioncall{tax_agg}(df, rank = \hlstring{'family'}, db = \hlstring{'ncbi'})
agg
\end{alltt}
\begin{verbatim}
